Tropical mathematics for the Lambda-calculus

### 2023-07-26 CSL24/section3C.tex
Commit giornaliero
Commit notturno

%\begin{remark}\label{rmk:ModelsOfBSTLC}
%Similarly to It is known {\color{red}(([reference??] e dire meglio)} 
For $\BSTLC$, one can see that the exponential $!$ can be ``decomposed'' into a family of \emph{graded} exponentials functors $!_n:\LREL\to\LREL$ ($n\in\BB N$), defined by taking multisets of cardinality \emph{at most} $n$. %  lift to functors 
The sequence $(!_n)_{n\in\N}$ gives rise to a $\N$-graded linear exponential comonad on (the SMC) $\LREL$, %satisfying the adjunction: $\LREL(Z\otimes !_{n}X,Y) \simeq \LREL(Z, !_{n}X\multimap Y)$.
i.e.\ $(\LREL,(!_n)_{n\in\N})$ is a model of $\BSTLC$. 
In particular, bounded arrow types are interpreted via $\model{!_{n}A\multimap B}:= \C M_{\leq n}(\model A) \times \model B$ and so, 
if e.g.\ $\model o=\{\star\}$, then $\model{A}$ is a finite set for any type $A$ of $\BSTLC$.
%\end{remark}
%Bounded types are interesting because of the following proposition:
%\begin{proposition}
%For all bounded types $A,B$, the morphisms from $\model A$ to $\model B$ (in all parametric relational models) correspond to polynomials.
%\end{proposition}
%\begin{proof}
%It suffices to check that $\model A$ is finite for all bounded types $A$. Indeed this implies that a morphism $t:\model A\to \model B$ is a finite matrix $t: \model A \times \model B \to \Lawv$.Hence, its corresponding map $\widehat t:\Lawv^{\model A} \to \Lawv^{\model B}$ is a polynomial.
%\end{proof}

@@

For what concerns $\STDLC$, by patiently checking all the needed axioms {\color{red}or by using  LEMAY??}, one sees that
%\begin{proposition}[{\color{red}LEMAY??}]\label{thm:LREL!CCDC}
 $\LREL_!$ becomes a CC$\partial\lambda$C, i.e.\ a model of the $\STDLC$, when equipped with the \emph{tropical differential operator} $D:\HOM{\LREL}{!X}{Y}\to \HOM{\LREL}{!(X+X)}{Y}$ defined as $(Dt)_{\mu\oplus\rho,b}=t_{\rho+\mu,b}$ if $\#\mu=1$ and as $\infty$ otherwise (using the iso $(\mu,\rho)\in !Z\times !Z'\mapsto\mu\oplus\rho \in !(Z+Z')$).
%\end{proposition}
%\begin{remark}For $t\in\HOM{\LREL}{!X}{Y}$, we have: $D^2 t\in\HOM{\LREL_!}{(X+X)+(X+X)}{Y}$, where $(D^2 t)_{(\rho\oplus\rho')\oplus(\nu\oplus\nu'),b}$ equals $t_{\nu+\nu'+\rho',b}$ if $\rho=\emptyset$ and $\#\rho'=1=\#\nu$; it equals $t_{\rho+\nu',b}$ if $\rho'=\emptyset=\nu$ and $\#\rho=1$; it equals $\infty$ otherwise.\end{remark}
%This ensures that one can define a sound interpretation of $\STDLC$-terms in the standard way (see [Section 4.3, \cite{Manzo2010}]).
This model is also well-behaved w.r.t.\ the Taylor expansion, as expressed by the following two results.
As it can be patientely checked, in $(\LREL_!,D)$ all morphisms can be Taylor expanded  (see \cite[Definition 4.22]{Manzo2012}):

\begin{theorem}\label{thm:modelsTaylor}
 For all $t\in\HOM{\LREL_!}{Z}{X\multimap Y}$, $s\in\HOM{\LREL_!}{Z}{X}$ we have:%, the evaluation of $t$ over $s$ yields 
 \begin{align}\label{eq:taylorcat}
  \RM{ev}\circ_!\langle t,s\rangle =
  \inf\limits_{n\in\N}
  \set{((\dots((\Lambda^- t)\star s)\star \dots)\star s)\circ_! \langle \RM{id},\infty \rangle}.
 \end{align} 
\end{theorem}
%It is worth discussing the formula above a bit more. 
Here,
%:\HOM{\LREL}{!Z}{X\multimap Y}\to \HOM{\LREL}{!(Z+X)}{Y}$
%$\star:\HOM{\LREL}{!(Z+X)}{Y}\times\HOM{\LREL}{!Z}{X}\to \HOM{\LREL}{!(Z+X)}{Y}$ is defined as 
$u\star s= (Du)\circ_{!} \langle \langle  \infty, s\circ_{!} \pi_{1}\rangle,\mathrm{id}\rangle$ corresponds to the application of the derivative of $u$ on $s$, and $\Lambda^-$ is the uncurry operator.
Hence the right-hand term in \eqref{eq:taylorcat} corresponds to the $\inf$ of the $n$-th derivative of $\Lambda^{-}t$ applied to ``$n$ copies'' of $s$,  i.e.~it coincides with the tropical %interpretation of the
version of the usual Taylor expansion.
%Moreover, since $\LREL_!$ has countable sums (all $\inf$'s converge), and thanks to equation \eqref{eq:taylorcat}, an immediate adaptation of the proof of [Theorem 4.23, \cite{Manzo2012}] entails that the interpretation of the $\STDLC$-Taylor expansion of a $\STLC$-term $M$ given in \eqref{eq:taylor}, converges to the interpretation of $M$.
Finally, since $\LREL_!$ has countable sums (all countable $\inf$s converge), an immediate adaptation of the proof of [Theorem 4.23, \cite{Manzo2012}] shows:

@@

The fact that the coefficients are either $0$ or $\infty$ can be seen as the fact that $\STLC$ is relatively trivial, from this point of view.
As we will discuss in \autoref{sec:app}, non-trivial coefficients appear when interpreting more interesting languages, like probabilistic ones.
However, in this paper we shall still focus on $\STLC$, in order to set the basis of the further studies and in order to see some already interesting properties of the interpretation.

>>>>

%\begin{remark}\label{rmk:ModelsOfBSTLC}
%Similarly to It is known {\color{red}(([reference??] e dire meglio)} 
For $\BSTLC$, one can see that the exponential $!$ can be ``decomposed'' into a family of \emph{graded} exponentials functors $!_n:\LREL\to\LREL$ ($n\in\BB N$), defined on objects by taking multisets of cardinality \emph{at most} $n$. %  lift to functors 
The sequence $(!_n)_{n\in\N}$ gives rise to a $\N$-graded linear exponential comonad on (the SMC) $\LREL$, %satisfying the adjunction: $\LREL(Z\otimes !_{n}X,Y) \simeq \LREL(Z, !_{n}X\multimap Y)$.
i.e.\ $(\LREL,(!_n)_{n\in\N})$ is a model of $\BSTLC$. 
In particular, bounded arrow types are interpreted via $\model{!_{n}A\multimap B}:= !_{n}\model A \times \model B$ and so, 
if $\model *$ is finite, then $\model{A}$ is a finite set for any type $A$ of $\BSTLC$.
%\end{remark}
%Bounded types are interesting because of the following proposition:
%\begin{proposition}
%For all bounded types $A,B$, the morphisms from $\model A$ to $\model B$ (in all parametric relational models) correspond to polynomials.
%\end{proposition}
%\begin{proof}
%It suffices to check that $\model A$ is finite for all bounded types $A$. Indeed this implies that a morphism $t:\model A\to \model B$ is a finite matrix $t: \model A \times \model B \to \Lawv$.Hence, its corresponding map $\widehat t:\Lawv^{\model A} \to \Lawv^{\model B}$ is a polynomial.
%\end{proof}

@@

For what concerns $\STDLC$, applying \cite[Theorem 6.1]{lemay2020} one can check that
%\begin{proposition}[{\color{red}LEMAY??}]\label{thm:LREL!CCDC}
 $\LREL_!$ becomes a CC$\partial\lambda$C, i.e.\ a model of the $\STDLC$, when equipped with the \emph{tropical differential operator} $D:\HOM{\LREL}{!X}{Y}\to \HOM{\LREL}{!(X\& X)}{Y}$ defined as $(Dt)_{\mu\oplus\rho,b}=t_{\rho+\mu,b}$ if $\#\mu=1$ and as $\infty$ otherwise (using the iso $(\mu,\rho)\in !Z\times !Z'\mapsto\mu\oplus\rho \in !(Z+Z')$).
%\end{proposition}
%\begin{remark}For $t\in\HOM{\LREL}{!X}{Y}$, we have: $D^2 t\in\HOM{\LREL_!}{(X+X)+(X+X)}{Y}$, where $(D^2 t)_{(\rho\oplus\rho')\oplus(\nu\oplus\nu'),b}$ equals $t_{\nu+\nu'+\rho',b}$ if $\rho=\emptyset$ and $\#\rho'=1=\#\nu$; it equals $t_{\rho+\nu',b}$ if $\rho'=\emptyset=\nu$ and $\#\rho=1$; it equals $\infty$ otherwise.\end{remark}
%This ensures that one can define a sound interpretation of $\STDLC$-terms in the standard way (see [Section 4.3, \cite{Manzo2010}]).
This model is also well-behaved w.r.t.\ the Taylor expansion, as expressed by the following two results.
As it can be patientely checked, in $(\LREL_!,D)$ all morphisms can be Taylor expanded  (see \cite[Definition 4.22]{Manzo2012}):

\begin{theorem}\label{thm:modelsTaylor}
 For all $t\in\HOM{\LREL_!}{Z}{X\multimap Y}$, $s\in\HOM{\LREL_!}{Z}{X}$ we have:%, the evaluation of $t$ over $s$ yields 
 \begin{align}\label{eq:taylorcat}
  \RM{ev}\circ_!\langle t,s\rangle =
  \inf\limits_{n\in\N}
  \set{((\dots((\Lambda^- t)\star s)\star \dots)\star s)\circ_! \langle \RM{id},\infty \rangle}.
 \end{align} 
\end{theorem}
%It is worth discussing the formula above a bit more. 
Here,
%:\HOM{\LREL}{!Z}{X\multimap Y}\to \HOM{\LREL}{!(Z+X)}{Y}$
%$\star:\HOM{\LREL}{!(Z+X)}{Y}\times\HOM{\LREL}{!Z}{X}\to \HOM{\LREL}{!(Z+X)}{Y}$ is defined as 
$u\star s= (Du)\circ_{!} \langle \langle  \infty, s\circ_{!} \pi_{1}\rangle,\mathrm{id}\rangle$ corresponds to the application of the derivative of $u$ on $s$, and $\Lambda^-$ is the uncurry operator.
Hence the right-hand term in \eqref{eq:taylorcat} corresponds to the $\inf$ of the $n$-th derivative of $\Lambda^{-}t$ applied to ``$n$ copies'' of $s$,  i.e.~it coincides with the tropical %interpretation of the
version of the usual Taylor expansion.
%Moreover, since $\LREL_!$ has countable sums (all $\inf$'s converge), and thanks to equation \eqref{eq:taylorcat}, an immediate adaptation of the proof of [Theorem 4.23, \cite{Manzo2012}] entails that the interpretation of the $\STDLC$-Taylor expansion of a $\STLC$-term $M$ given in \eqref{eq:taylor}, converges to the interpretation of $M$.
Finally, since $\LREL_!$ has countable sums (all countable $\inf$s converge), an immediate adaptation of the proof of \cite[Theorem 4.23]{Manzo2012} shows:

@@

The fact that the coefficients are either $0$ or $\infty$ can be seen as the fact that $\STLC$ is relatively trivial, from this point of view.
As we will discuss in \autoref{sec:app}, non-trivial coefficients appear when interpreting more interesting languages, like probabilistic ones.
However, in this paper we shall still focus on $\STLC$, in order to set the basis of the further studies and in order to see some interesting properties already of its interpretation.

### 2023-07-26 CSL24/section4.tex
Commit notturno

For starting, remark that although a matrix $t\in\HOM{\LREL_!}{X}{Y}$ yields a linear map $\Lawv^{!X}\to\Lawv^Y$, by exploiting the coKleisli structure we can also ``express it in the base $X$'', i.e.\ see it as a \emph{non-linear} map $t^!:\Lawv^X\to\Lawv^Y$, by setting $t^!(x):=t\circ_! x$. %(we are identifying $\Lawv^X$ with the set $\HOM{\LREL_!}{\emptyset}{X}$ of the \emph{points} of $X$).
Concretely, we have $t^!(x)_b=\inf_{\mu\in !X} \set{\mu x+ t_{\mu,b}}$, where $\mu x:=\sum_{a\in X} \mu(a)x_a$.
In this section we establish some topological and metric properties of such functions, this way highlighting the rich and interesting topological and metric structure of the category $\LREL_{!}$.
In analogy with \cite{DanEhrh2011}, we could call them \emph{tropical entire functions}, but since, as we explain in a moment, those are generalised \emph{tropical Laurent series}, we will use the latter terminology, or \emph{tLs} for short.
The literature in tropical mathematics is often recent (e.g.~\cite{Porzio2021}), and several results we prove in this section are, to our knowledge, new.

@@

@@

 In a similar way, tLs
$f:\Lawv^X\to\Lawv^Y$ 
 with \emph{finite} supports $\C F_b=\set{\mu\in!X\mid\matr f_{\mu,b}\neq\infty}$, and which have thus shape $f(x)_b=\min_{\mu\in\C F} \set{\mu x+ t_{\mu,b}}$, are generalisation of usual tropical polynomials to the case of possibly infinitely many variables.
For instance, for a $t\in\HOM{\LREL}{!_n X}{Y}$ (this is in particular the case for the interpretations of $\BSTLC$-terms), we will consider the (generalised) tropical polynomial $t^!:\Lawv^X\to\Lawv^Y$ defined by $t^!(x)_b:=\inf_{\mu\in !_nX}\{\mu x+t_{\mu,b}\}$.
% 
% , i.e.~for a \emph{finite} set $\C F\subseteq \, !X$.
%%Remark that we also find usual \emph{tropical polynomials} of tropical geometry as a particular case: they 
%correspond to the tLs for which the support $\set{n\in\N\mid\widehat f(n)\neq\infty}$ of $\widehat f$ is \emph{finite}. 
%Actually, for us \emph{tropical polynomial} will mean a function 
%\end{remark}

@@

\begin{theorem}\label{thm:cont}
 All tLs $f:\Lawv^X\to\Lawv$ are continuous on $(0,\infty)^X$, w.r.t.\ to the norm $\norm{\cdot}_\infty$.
\end{theorem}
\begin{proof}
%It follows after adapting \cite[Proposition 4.4]{Cobzas2017} in order to prove that if a real-valued function on a locally convex topological $\BB R$-vector space is, locally around $x$, concave and bounded by a finite constant, then it is continuous at $x$.
\cite[Proposition 4.4.(3)]{Cobzas2017} shows that if a real-valued convex function with domain a convex open subset of a locally convex topological $\BB R$-vector space is, locally around any point, bounded from above by a finite non-zero constant, then it is continuous on all its domain.
Now, our $f$ is concave, so $-f$ is convex.
Also, since $f\geq 0$ on all $\Lawv^X$, we have e.g.\ $-f\leq 1$ on $\Lawv^X$.
Remark that we cannot immediately apply \cite[Proposition 4.4.(3)]{Cobzas2017}, since $\Lawv^X$ is not even a $\BB R$-vector space.
But $(0,\infty)^X\subseteq \BB R^X$ is open and convex, and $\BB R^X$ is a locally convex topological $\BB R$-vector space.
Thus \cite[Proposition 4.4.(3)]{Cobzas2017} entails the continuity of $-f\big|_{(0,\infty)^X}$, hence that of $f\big|_{(0,\infty)^X}$.
\end{proof}

@@

\begin{corollary}
Let $\model A$ be a finite set.
\begin{enumerate}
\item $\model{\Gamma \vdash_{\BSTLC} M:A}^!:\Lawv^{\model\Gamma} \to \Lawv^{\model A}$ is a \emph{tropical polynomial}, thus (as $\model A$ is finite), a \emph{Lipschitz} function.
\item $\model{\Gamma \vdash_{\STLC} M:A}^!:\Lawv^{\model\Gamma} \to \Lawv^{\model A}$ is a \emph{locally} Lipschitz map.
\item $\Te{M}$ decomposes $\model{\Gamma \vdash_{\STLC} M:A}^!$ as an $\inf_{t\in\Te{M}}\model{\Gamma\vdash_{\STDLC} t:A}^!$ of \emph{tropical polynomials}, thus (as $\model A$ is finite), \emph{Lipschitz} functions.
\end{enumerate}
\end{corollary} 
\begin{proof}
1). We already observed that the interpretation of a $\BSTLC$-type $B$ is alway a finite set, so thus is $!_n \model B$.
So the $\inf$ defining $(\_)^!$ is actually a $\min$.
So the interpretation of a bounded term is a tropical polynomial, so we apply Corollary~\ref{prop:polylip} to each coordinate of the image, and since $\model A$ is finite (by taking the maximum Lipschitz constant among the finite number of them), we obtain the thesis.
2). It follows immediately from Theorem~\ref{thmTLSlocLip} and the fact that $\model A$ is finite.
3). It follows from \autoref{cor:T(M)=M} plus the easily checked fact that, for $(f_n)_{n\in\N}\subseteq\Lawv^{!X\times Y}$, we have $\left(\inf_{n\in\N} f_n\right)^!:\Lawv^X\to\Lawv^Y$, with $\left(\inf_{n\in\N} f_n\right)^!=\inf_{n\in\N} f_n^!$.
\end{proof}

@@

A consequence of (3) is that the distance between two programs can always be approximated via Lipschitz tropical polynomial approximants of the initial two programs.
\begin{corollary}
 Let $\Gamma \vdash_{\STLC} M:A$ and $\Delta\vdash_{\STLC} N:B$.
 For all $\epsilon>0$, $x\in\Lawv^{\model \Gamma}$, $b\in\model A$, there exist $t\in\Te{M}$, $u\in\Te{N}$ s.t.\ $\big| \model{\Gamma \vdash_{\STLC} M:A}^!(x)_b - \model{\Delta \vdash_{\STLC} N:B}^!(x)_b \big| \leq 2\epsilon + \big| \model{\Gamma \vdash_{\STDLC} t:A}^!(x)_b - \model{\Delta \vdash_{\STDLC} u:B}^!(x)_b \big|$.
\end{corollary}

>>>>

For starting, remark that although a matrix $t\in\HOM{\LREL_!}{X}{Y}$ yields a linear map $\Lawv^{!X}\to\Lawv^Y$, by exploiting the coKleisli structure we can also ``express it in the base $X$'', i.e.\ see it as a \emph{non-linear} map $t^!:\Lawv^X\to\Lawv^Y$, by setting $t^!(x):=t\circ_! x$. %(we are identifying $\Lawv^X$ with the set $\HOM{\LREL_!}{\emptyset}{X}$ of the \emph{points} of $X$).
Concretely, we have $t^!(x)_b=\inf_{\mu\in !X} \set{\mu x+ t_{\mu,b}}$, where $\mu x:=\sum_{a\in X} \mu(a)x_a$.

In this section we establish some topological and metric properties of such functions, this way highlighting the rich and interesting topological and metric structure of the category $\LREL_{!}$.
In analogy with \cite{DanEhrh2011}, we could call them \emph{tropical entire functions}, but since, as we explain in a moment, those are generalised \emph{tropical Laurent series}, we will use the latter terminology, or \emph{tLs} for short.
The literature in tropical mathematics is often recent (e.g.~\cite{Porzio2021}), and several results we prove in this section are, to our knowledge, new.

@@

\begin{remark}
 Since $\model{(x_i:A_i)_{i=1}^n\vdash_{\STLC} M:B}\in\HOM{\LREL_!}{\mathlarger{\&}_{i=1}^n \model{A_i}}{\model B}\simeq\Lawv^{\prod_{i=1}^n ! \model {A_i}\times \model B}$, the tLs $\model{(x_i:A_i)_{i=1}^n\vdash_{\STLC} M:B}^!:\Lawv^{\mathlarger{+}_{i=1}^n \model{A_i}}\to\Lawv^{\model B} \simeq \prod_{i=1}^n\Lawv^{\model{A_i}}\to\Lawv^{\model B}$ is defined by $t^!(x^1,\dots,x^n)_b:=\inf_{\mu_i\in ! \model{A_i}}\left\{\sum_{i=1}^n\mu_i x^i+t_{(\mu_1,\dots,\mu_n),b}\right\}$.
Same for $\STDLC$.
Finally, the interpretations of $\BSTLC$-terms are matrices $t\in\HOM{\LREL}{\bigotimes_{i=1}^n !_{n_i}X_i}{Y}=\Lawv^{\prod_{i=1}^n !_{n_i}X_i\times Y}$.
In such situation, we define $t^!:\Lawv^{\mathlarger{+}_{i=1}^n X_{i}}\to\Lawv^Y \simeq \prod_{i=1}^n\Lawv^{X_{i}}\to\Lawv^Y$ as $t^!(x^1,\dots,x^n)_b:=\inf_{\mu_i\in !_{n_i} X_i}\left\{\sum_{i=1}^n\mu_i x^i+t_{(\mu_1,\dots,\mu_n),b}\right\}$.
Clearly $t^!={\widetilde t}^!$, where $\widetilde t\in\Lawv^{\prod_{i=1}^n ! {X_i}\times Y}$ is the matrix $\widetilde t_{(\mu_1,\dots,\mu_n),b}:=t_{(\mu_1,\dots,\mu_n),b}$ if $\mu_i\in !_{n_i} X_i$ for all $i$, and $:=\infty$ otherwise, which has the same support as $t$.
So, such $t^!$'s are special cases of tLs and thus, \emph{Wlog}, in the following we only consider tLs.
\end{remark}

@@

 In a similar way, tLs
$f:\Lawv^X\to\Lawv^Y$ 
 with \emph{finite} supports $\C F_b=\set{\mu\in!X\mid\matr f_{\mu,b}\neq\infty}$, and which have thus shape $f(x)_b=\min_{\mu\in\C F} \set{\mu x+ t_{\mu,b}}$, are generalisation of usual tropical polynomials to the case of possibly infinitely many variables.
If $\model *$ is finite, this is in particular the case for the interpretations of $\BSTLC$-terms, since then the interpretation of any $\BSTLC$-type $B$ is a finite set, and thus also $!_n \model B$ is.
% 
% , i.e.~for a \emph{finite} set $\C F\subseteq \, !X$.
%%Remark that we also find usual \emph{tropical polynomials} of tropical geometry as a particular case: they 
%correspond to the tLs for which the support $\set{n\in\N\mid\widehat f(n)\neq\infty}$ of $\widehat f$ is \emph{finite}. 
%Actually, for us \emph{tropical polynomial} will mean a function 
%\end{remark}

@@

\begin{theorem}\label{thm:cont}
 All tLs $f:\Lawv^X\to\Lawv$ are continuous on $(0,\infty)^X$, w.r.t.\ to the norm $\norm{\cdot}_\infty$.
\end{theorem}
\begin{proof}
%It follows after adapting \cite[Proposition 4.4]{Cobzas2017} in order to prove that if a real-valued function on a locally convex topological $\BB R$-vector space is, locally around $x$, concave and bounded by a finite constant, then it is continuous at $x$.
\cite[Proposition 4.4.(3)]{Cobzas2017} shows that if a real-valued convex function with domain a convex open subset of a locally convex topological $\BB R$-vector space is, locally around any point, bounded from above by a finite non-zero constant, then it is continuous on all its domain.
Now, our $f$ is concave, so $-f$ is convex.
Also, since $f\geq 0$ on all $\Lawv^X$, we have e.g.\ $-f\leq 1$ on $\Lawv^X$.
%Remark that we cannot immediately apply \cite[Proposition 4.4.(3)]{Cobzas2017}, since $\Lawv^X$ is not even a $\BB R$-vector space.
Now $(0,\infty)^X\subseteq \BB R^X$ is open and convex, and {\color{red}$\BB R^X$ is a locally convex topological $\BB R$-vector space}, so \cite[Proposition 4.4.(3)]{Cobzas2017} entails the continuity of $-f\big|_{(0,\infty)^X}$, hence that of $f\big|_{(0,\infty)^X}$.
\end{proof}

@@

\begin{corollary}
Let $\model A$ be a finite set.
\begin{enumerate}
\item $\model{\Gamma \vdash_{\BSTLC} M:A}^!:\Lawv^{\model\Gamma} \to \Lawv^{\model A}$ is a \emph{tropical polynomial}, thus (as $\model A$ is finite), a \emph{Lipschitz} function.
\item $\model{\Gamma \vdash_{\STLC} M:A}^!:\Lawv^{\model\Gamma} \to \Lawv^{\model A}$ is a \emph{locally} Lipschitz map.
\item $\Te{M}$ decomposes $\model{\Gamma \vdash_{\STLC} M:A}^!$ as an $\inf_{t\in\Te{M}}\model{\Gamma\vdash_{\STDLC} t:A}^!$ of \emph{tropical polynomials}, thus (as $\model A$ is finite), \emph{Lipschitz} functions.
\end{enumerate}
\end{corollary} 
\begin{proof}
1). Since $\model A$ is finite, also $\model *$ is.
Thus, as we already observed, the interpretation of a bounded term is a tropical polynomial.
Now we apply Corollary~\ref{prop:polylip} to each coordinate of the image, and by taking the maximum Lipschitz constant among the finite number $\mathrm{Card}(\model A)$ of them, we obtain the thesis.
2). It follows immediately from Theorem~\ref{thmTLSlocLip} and the fact that $\model A$ is finite.
3). It follows from \autoref{cor:T(M)=M} plus the easily checked fact that, for $(f_n)_{n\in\N}\subseteq\Lawv^{!X\times Y}$, we have $\left(\inf_{n\in\N} f_n\right)^!:\Lawv^X\to\Lawv^Y$, with $\left(\inf_{n\in\N} f_n\right)^!=\inf_{n\in\N} f_n^!$.
\end{proof}

@@

A consequence of (3) is that the pointwise distance between two interpretations of programs can always be bounded via Lipschitz tropical polynomial approximants of the initial two programs.
\begin{corollary}
 Let $\Gamma \vdash_{\STLC} M:A$ and $\Delta\vdash_{\STLC} N:B$.
 For all $\epsilon>0$, $x\in\Lawv^{\model \Gamma}$, $b\in\model A$, there exist $t\in\Te{M}$, $u\in\Te{N}$ s.t.\ $\big| \model{\Gamma \vdash_{\STLC} M:A}^!(x)_b - \model{\Delta \vdash_{\STLC} N:B}^!(x)_b \big| \leq 2\epsilon + \big| \model{\Gamma \vdash_{\STDLC} t:A}^!(x)_b - \model{\Delta \vdash_{\STDLC} u:B}^!(x)_b \big|$.
\end{corollary}

### 2023-07-26 CSL24/section6.tex
Commit notturno

>>>>

\begin{remark}
 $\QREL$ is (equivalent to) a subcategory of the category $Q\SF{Mod}$ of \emph{complete} $Q$-semimodules.
 If $\QREL$ corresponds to considering semimodules (the $Q^X$'s) whose vectors are given in coordinates w.r.t.\ a \emph{fixed base} (the set $X$), $Q\SF{Mod}$ corresponds to considering semimodules in abstract, without fixing a base.
\end{remark}
